Kronos \& Krios: Evidence for accretion of a massive, rocky planetary system in a comoving pair of solar-type stars

### 2017-08-28 paper/ms.tex
more lit summary and abstract edit

\begin{abstract}
  We report and discuss the discovery of a comoving pair of bright solar-type
  stars, HD~240430 and HD~240429, with a significant difference in their
  chemical abundances.
  The two stars have an estimated separation of $\approx 0.6$~pc at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities,
  %($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence),
  as inferred from \gaia\ \tgas\ parallaxes and proper motions, and
  high-precision radial velocity measurements.
  Stellar parameters determined from high-resolution Keck HIRES spectra
  indicate that both stars are $\sim 4$~Gyr old,
  yet both stars show very high surface \elem{Li} abundance for this age.
  The more metal-rich of the two, HD~240430, shows an enhancement of refractory
  ($\Tcondens>1200$~K) elements by $\approx 0.2$~dex but not as large
  an enhancement of (moderately) volatile elements ($\Tcondens<1200$~K; \elem{C},
  \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}).
  This is the largest and the most significant chemical difference found in
  a wide binary pair yet.
  The proximity in phase-space and ages between the two stars suggests that
  they formed together with the same composition, at odds with the observed
  differences in metallicity and abundance patterns.
  We therefore suggest that the star HD~240430, ``Kronos'', recently accreted
  $\approx 20$~\mearth\ of rocky material after birth selectively enhancing the
  refractory elements in its surface and convective envelope.
  The other star HD~240429, ``Krios'', may also have had a similar accretion
  event given its high surface \elem{Li} abundance but of much smaller amount
  ($\approx 4$~\mearth).
  %Follow-ups to study the existence and properties of the planetary systems
  %around these stars will help illuminate planet-star interactions.
\end{abstract}

@@

\todo{might as well note no orbits ever established at these separations}
\todo{smoh: discussion of HD~1333131AB and 16 Cyg is missing, but will be added}

@@

{\bf WASP-94A/WASP-94B:}
Each star in a pair of F8 and F9 stars with super-solar metallicity
($\feh\approx 0.3$) hosts a hot Jupiter.
The planet around WASP-94A is transiting with a misaligned, probably retrograde
circular ($e<0.13$) orbit, while that hosted by WASP-94B is a little more
massive by $\sim 0.15$~\mjupiter\ and closer in, aligned with the host star.
WASP-94A shows a depletion of $0.02$~dex in volatile and moderately volatile
elements ($\Tcondens < 1200$~K) and an enhancement of $0.01$~dex in refractory
elements ($\Tcondens>1200$~K) relative to WASP-94B, with a claimed median
uncertainty of $0.006$~dex among all elements
resulting in a statistically significant non-zero slope between
$\Delta\elemH{X}$ and $\Tcondens$ (\citealt{Teske:2016aa}).\footnote{
  %TODO: double check the following statement
  Note that the condensation
  temperature $\Tcondens$ used is for solar system composition
  gas, which can differ from that of higher metallicity gas.}
Multiple possibilities related to the formation and evolution
of planetary systems around each star as well as causes unrelated to planets
such as dust cleansing during the fully convective phase or different rotation
and granulation between the stars were considered, but none was favored.

>>>>

\begin{abstract}
  We report and discuss the discovery of a comoving pair of bright solar-type
  stars, HD~240430 and HD~240429, with a significant difference in their
  chemical abundances.
  The two stars have an estimated separation of $\approx 0.6$~pc at a distance of
  $r\approx 100~\pc$ with nearly identical three-dimensional velocities,
  %($|\Delta \bs{v}| < XX~\kms$ with 95\% confidence),
  as inferred from \gaia\ \tgas\ parallaxes and proper motions, and
  high-precision radial velocity measurements.
  Stellar parameters determined from high-resolution Keck HIRES spectra
  indicate that both stars are $\sim 4$~Gyr old.
  The more metal-rich of the two, HD~240430, shows an enhancement of refractory
  ($\Tcondens>1200$~K) elements by $\approx 0.2$~dex but not as large
  an enhancement of (moderately) volatile elements ($\Tcondens<1200$~K; \elem{C},
  \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}).
  This is the largest and the most significant chemical difference found in
  a wide binary pair yet.
  Additionally, it shows anomalously high surface lithium abundance
  ($A(\elem{Li}=2.75$), higher by $0.5$~dex than its companion.
  The proximity in phase-space and ages between the two stars suggests that
  they formed together with the same composition, at odds with the observed
  differences in metallicity and abundance patterns.
  We therefore suggest that the star HD~240430, ``Kronos'', recently accreted
  $\approx 20$~\mearth\ of rocky material after birth selectively enhancing the
  refractory elements in its surface and convective envelope.
\end{abstract}

@@

\todo{smoh: discussion of HD~1333131AB and 16 Cyg is missing, but will be added}

{\bf 16 Cygni A/B:}
The chemical composition of this well know pair of solar-type stars (G1.5/G3)
has been studied many times.
The hotter star 16 Cyg A has no detected planets, but a M dwarf companion $\sim
70$~AU away in projected separation which is probably physically associated
(\citealt{2002ApJ...581..654P}), and may have affected planet formation process
around the star (\citealt{1996ApJ...458..312J,2005MNRAS.363..641M}).
The other star, 16 Cyg B, hosts a giant planet on an eccentric orbit ($e=0.63$,
\citealt{1997ApJ...483..457C}).
While past measurements of metallicity and abundance difference between the two stars
claimed conflicting results (\citealt{2001ApJ...553..405L,2011ApJ...737L..32S}),
recent studies using high quality spectra consistently reported that
A is more metal rich than B by $\approx 0.04 \pm 0.005$~dex.
However, there is still a disagreement between studies on
whether abundance differences shows a correlation with $\Tcondens$ as well as
its interpretation.

@@

{\bf WASP-94A/B:}
Each star in a pair of F8 and F9 stars with super-solar metallicity
($\feh\approx 0.3$) hosts a hot Jupiter.
The planet around WASP-94A is transiting with a misaligned, probably retrograde
circular ($e<0.13$) orbit, while that hosted by WASP-94B is a little more
massive by $\sim 0.15$~\mjupiter\ and closer in, aligned with the host star.
WASP-94A shows a depletion of $0.02$~dex in volatile and moderately volatile
elements ($\Tcondens < 1200$~K) and an enhancement of $0.01$~dex in refractory
elements ($\Tcondens>1200$~K) relative to WASP-94B, with a claimed median
uncertainty of $0.006$~dex among all elements
resulting in a statistically significant non-zero slope between
$\Delta\elemH{X}$ and $\Tcondens$ (\citealt{Teske:2016aa}).\footnote{
  %TODO: double check the following statement
  Note that the condensation
  temperature $\Tcondens$ used is for solar system composition
  gas, which can differ from that of higher metallicity gas.}
Multiple possibilities related to the formation and evolution
of planetary systems around each star as well as causes unrelated to planets
such as dust cleansing during the fully convective phase or different rotation
and granulation between the stars were considered, but none was favored.

{\bf $\zeta^1/\zeta^2$ Reticuli (HD 20807/HD 20766):}
With a projected separation of $\approx 3700$~AU, both solar-type stars in this pair have
no detected planets.
However, $\zeta^2$ hosts a debris disk detected via infrared excess
(\citealt{2008ApJ...674.1086T}) as well as direct imaging
(\citealt{2010A&A...518L.131E}).
Both stars have super-solar metallicity of $\approx 0.2$~dex.
A differential abundance analysis using high-resolution spectra
shows that $\zeta^1$ is more metal rich than $\zeta^2$ by $\sim 0.02 \pm 0.003$~dex,
and that there is a positive slope between the abundance differences of 24 species
and $\Tcondens$.
A possible explanation proposed is that the relative lack of refractory elements
in $\zeta^2$ is because they are locked up in rocky bodies
that make up its debris disk (\citealt{2016A&A...588A..81S}).

{\bf HD~133131A/B:}
